# Data Cleaning {#clean}

After the initial parsing of the data in the acquisition phase, there was not a significant amount of cleaning to be done. The cleaning stage primarily consisted of organizing the variables in a way that was intuitive to work with. We chose to handle NA's by replacing them with the group mean determined by their Carnegie classification.
For the demographics data set, we first dropped extraneous columns. These were columns which consisted of more than 90% NA values. Then, we renamed the variables so that they would be more intuitive to a general audience (and so that we could remember them better). There were four schools within the data that did not report any demographic variables for every year that they were in the data set. These schools were removed using drop_na(black), because the 'black' column did not possess any NAs other than for these four schools.

For both data sets, we chose to recode the Carnegie Classification variable (called r_status) as R1 and R2 to increase the interpretability of the data. These are converted to factors in the analysis section so that they can be input into the models.

## Demographics

```{r, message=FALSE}

#IMPORTING PACKAGES

library(tidyverse)

```


```{r, message = FALSE}

# READING IN THE CSV

dem <- read.csv("Proj1Data/Demographics15to19.csv")

# CLEANING THE DATA FRAME

dem_clean <- dem[,-c(1, 17, 20:24)] %>%
  rename("university" = instnm,
  "r_status" = ccbasic, 
  "state" = stabbr, 
  "white" = ugds_white, 
  "black" = ugds_black, 
  "hispanic" = ugds_hisp, 
  "asian" = ugds_asian, 
  "indigenous" = ugds_aian, 
  "nhpi" = ugds_nhpi, 
  "nra" = ugds_nra, 
  "unknown" = ugds_unkn) %>% 
  drop_na(black)

# RECODING THE CARNEGIE CLASSIFICATION (R_STATUS) VARIABLE

dem_clean$r_status <- recode(dem_clean$r_status, `15` = 'R1', `16` = "R2")

# NA HANDLING

dem_clean <- dem_clean %>%  
  group_by(r_status) %>%
  mutate_at(c("female", "first_gen", "poverty_rate", "veteran", "unemp_rate"), funs(ifelse(is.na(.), mean(., na.rm = TRUE),.)))

# WRITING TO A NEW CSV

write.csv(dem_clean, "Proj1Data/cleanDemographics1519.csv")

```

## Financials 

The Financials data set was easier to work with; as such, we simply renamed the common variables that we had renamed with demographics and re-coded the Carnegie Classification.

```{r, message= FALSE}

# READING IN THE CSV

fin <- read.csv('Proj1Data/Financials15to19.csv')

# RENAMING VARIABLES SO THAT THEY MATCH THE DEMOGRAPHIC DATA FRAME

fin_clean <- fin[,-c(2)] %>%
  rename("university" = instnm,
  "r_status" = ccbasic, 
  "state" = stabbr)

# RECODING THE CARNEGIE CLASSIFICATION (R_STATUS) VARIABLE

fin_clean$r_status <- recode(fin_clean$r_status, `15` = 'R1', `16` = "R2")

# WRITING TO A NEW CSV

write.csv(fin_clean, "Proj1Data/cleanFinancials1519.csv")

```

```{r}

#DISPLAYING THE NEW CSV

head(dem_clean, 5) 

```

