# Methods and Results {#analysis}

## Methods 

**Sample.**  
: The sample for this study consisted of universities in the United States of America. In this way, we were able to examine what was nearly the population of college students in America. However, the high frequency of NA variables in some categories hindered 

**Procedure.**
: The data was collected by accessing the public College Scorecard API. The procedure for extracting and cleaning the data is explained in the Data Extraction and Data Clean sections. 

**Analysis Strategy.**
: We quantified student outcomes using the median debt and unemployment rate variables. These variables were selected because they are tangibly related to financial outcomes. Other potential measures of success in the data, such as completion rates and death rates, contained either very sparse data or were ambiguous. For instance, neither transferring nor withdrawing from a University are necessarily indicative of a poor outcome (insofar as it relates to student success after attending the University). Both debt and unemployment, however, are. High rates of debt implicate a school with insufficient financial aid programs. High debt medians also potentially imply student’s are not placed into high-paying jobs that would allow them to expediently repay their debt. High rates of unemployment suggest that students have trouble finding jobs after graduation. As the goal of a university education is largely to educate a student to assist with their career, students being unable to find jobs is a poor outcome. 

The demographic variables were selected by inspecting all the demographic variables in the dataset and selecting the ones that were mostly complete. We decided to look at the demographic makeup of R1 and R2 universities to determine if there was a significant difference between the student body profiles. Race and gender are often discussed in relation to student success, due to a history of systematic discrimination within the American university system. These were our primary variables of interest. Specifically, we were interested in the ratio of White students to minority students. However, we briefly examined veteran and first generation status. 


## Results

```{r, message=FALSE}

library(moments)
library(dplyr)
library(tidyr)
library(stats)
library(sjstats)
library(ggplot2)
library(plotly)
library(tidyverse)
library(purrr)
library(RColorBrewer)
library(tmap)

```

```{r}
#READING IN THE CSVs

fin <- read.csv('Proj1Data/cleanFinancials1519.csv')
dem <- read.csv('Proj1Data/cleanDemographics1519.csv')
fin$r_status <- factor(fin$r_status, levels = c("R1", "R2"), labels = c(1,2))
dem$r_status <- factor(dem$r_status, levels = c("R1", "R2"), labels = c(1,2))

#TRANSFORMING THE DATASETS 

finR <- fin[fin$year == "latest",]
finR1 <- fin[fin$r_status == 1 & fin$year == "latest",]
finR2 <- fin[fin$r_status == 2 & fin$year == "latest",]
demR <- dem[dem$year == "latest",]
demR1 <- dem[dem$r_status == 1 & dem$year == "latest",]
demR2 <- dem[dem$r_status == 2 & dem$year == "latest",]
finR <- subset(finR, university %in% demR$university)
```

### Demographic Descriptive Statistics

**LOCATION OF R1'S vs. R2'S**

Before analyzing the demographic differences, we thought it was important to display how R1 and R2 schools are distributed across the United States. The demographic makeup of a school is potentially influenced by the demographics of in-state students which, at least at public institutions, are proportionally more likely to attend the school than students from any other state (see State Council of Higher Education for Virginia, 2019 for a case study of one state). Although we do not factor state and baseline state demographics into our model, it is a potential influential third variable that should be kept in mind while interpretting race and income demographics. 

```{r}

r1_state <- cdem %>%
  filter(r_status == "R1") %>%
  group_by(state) %>%
  distinct(university) %>%
  count()

r2_state <- cdem %>%
  filter(r_status == "R2") %>%
  group_by(state) %>%
  distinct(university) %>%
  count()

```

```{r}
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  lakecolor = toRGB('white')
)

r1_map <- plot_geo() %>%
  add_trace(
    z = ~r1_state$n,
    span = I(1),
    colorscale = 'Portland',
    zauto = F,
    zmax = 10,
    zmin = 1,
    locations = r1_state$state, 
    locationmode = 'USA-states'
  ) %>%
  colorbar(title = "# of Schools") %>%
  layout(geo = g, title = "Number of R1 Schools per State")

r2_map <- plot_geo() %>%
  add_trace(
    z = ~r2_state$n,
    span = I(1),
    colorscale = 'Portland',
    zauto = F,
    zmax = 10,
    zmin = 1,
    locations = r2_state$state, 
    locationmode = 'USA-states'
  ) %>%
  colorbar(title = "# of Schools") %>%
  layout(geo = g, title = "Number of R2 Schools per State")

r1_map
r2_map
```

**RACE**

The prelimary examination of the demographics data revealed that all of the race values, except for the proportion of white students, had a significant right skew. While the distributions are not normal, they are symmetric and bounded, which prevents extreme outliers. Below we provide a violin plot showing the distribution for each of the race variables grouped by their Carnegie classification. The mean is marked by a red bar, and the median is marked by a black bar. 

```{r, fig.height = 10}

dem_pal <- c("darkolivegreen", "darkolivegreen3", "dodgerblue4", "deepskyblue", "lavenderblush4", "lavenderblush2", "palevioletred1", "rosybrown1",  "tomato2",  "sienna 1", "slateblue3", "thistle1", "orange", "navajowhite2", "lightskyblue3", "lightsteelblue1")
  
race_demographics <- cdem %>%
  select(university, r_status, white, black, hispanic, asian, indigenous, nhpi, nra, unknown) %>%
  gather("race", "percentage", 3:10, -university)

race_demographics %>%
  ggplot(aes(x = factor(r_status), y = percentage, fill = interaction(r_status, race))) +
  geom_violin() + 
  stat_summary(fun=mean, geom="crossbar", linetype = 1, size=.2, color = "red") + 
  stat_summary(fun=median, geom="crossbar", linetype = 1, size=.2, color = "black") +
  labs(title = "Racial Demographics by Carnegie Classification", 
       subtitle = "Red line = mean | Black line = median\n",
       x = " ", 
       y = "% of Student Body") + 
  facet_wrap(~as.factor(race), nrow = 5) + 
  scale_fill_manual(values = dem_pal) + 
  theme_gray() + 
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        plot.title = element_text(hjust = .5), 
        legend.position = "none")

```


The five number summary for the 'white variable', along with the skewness and kurtosis, are reported below. 

```{r}

# DEMOGRAPHICS AT R1'S - WHITE STUDENTS 

summary(demR1$white)
skewness(demR1$white)
kurtosis(demR1$white)

```


```{r}

# DEMOGRAPHICS AT R2'S - WHITE STUDENTS 

summary(demR2$white)
skewness(demR2$white)
kurtosis(demR2$white)

```

**GENDER**

```{r, echo=FALSE}

cdem %>%
  filter(r_status == "R1", .preserve=TRUE) %>%
  select(female) %>%
  skewness()

cdem %>%
  filter(r_status == "R1", .preserve=TRUE) %>%
  select(female) %>%
  kurtosis()

cdem %>%
  filter(r_status == "R2", .preserve=TRUE) %>%
  select(female) %>%
  skewness()

cdem %>%
  filter(r_status == "R2", .preserve=TRUE) %>%
  select(female) %>%
  kurtosis()


```


```{r}

#Universities with the highest female/male ratio, the top 10 are mostly R2's for both categories

cdem %>%
  arrange(desc(female)) %>%
  filter(female > .50) %>%
  distinct(university, .keep_all = TRUE) %>%
  head(10)

cdem %>%
  arrange(female) %>%
  filter(female < .50) %>%
  distinct(university, .keep_all = TRUE) %>%
  head(10)

R1s <- cdem %>%
  filter(r_status == "R1")

ggplot(cdem, aes(x = female)) + 
  geom_histogram(bins = 40, color = "black", fill = "pink") + 
  labs(title = "Distribution of Female/Male Gender ratio at R1s")

R2s <- cdem %>%
  filter(r_status == "R2")

ggplot(R2s, aes(x = female)) + 
  geom_histogram(bins = 40, color = "black", fill = "skyblue") + 
  labs(title = "Distribution of Female/Male Gender ratio at R2s")

```


**OTHER DEMOGRAPHICS**

The non-racial variables (i.e., gender, first generation status, veteran status) did not posess the same skew. However, the mean was inflated due to how we coded NA's resulting in peculiar looking violin plots.  

```{r, fig.height = 10, fig.width = 6}

#Demographics for sparse columns 

other_demographics <- cdem %>%
  select(university, r_status, female, first_gen, veteran) %>%
  gather("demographic", "percentage", 3:5, -university)

other_plot <- other_demographics %>%
  ggplot(aes(x = factor(r_status), y = percentage, fill = interaction(r_status, demographic))) +
  geom_violin() + 
  stat_summary(fun=mean, geom="crossbar", linetype = 2, size=.1, color = "red") + 
  labs(title = "Other Demographic Factors by Carnegie Classification", 
       subtitle = "",
       x = " ", 
       y = "% of Student Body") + 
  facet_wrap(~as.factor(demographic), nrow = 5) + 
  scale_fill_brewer("Paired") + 
  theme_gray() + 
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), 
        plot.title = element_text(hjust = .5), 
        legend.position = "none")

other_plot

```


## Financial Descriptive Statistics

```{r}
# Cost
print('Tuition cost')
summary(finR1$costt4_a)
cost_hist1 <- ggplot(finR1, aes(x =costt4_a)) + 
  geom_histogram(bins = 40, color = "black", fill = "red") +
  labs(title = "Distribution of Tuition Costs at R1s", 
       subtitle = "", 
       x = "Cost (in American dollars)", 
       y = "Frequency") 
cost_hist1
```

```{r echo = FALSE}
summary(finR2$costt4_a)
cost_hist2 <- ggplot(finR2, aes(x =costt4_a)) + 
  geom_histogram(bins = 40, color = "black", fill = "pink") +
  labs(title = "Distribution of Tuition Costs at R2s", 
       subtitle = "", 
       x = "Cost (in American dollars)", 
       y = "Frequency") 
cost_hist2
```

```{r}
# Students debt
print('Student debt')
summary(finR1$grad_debt_mdn)
debt_hist1 <- ggplot(finR1, aes(x = grad_debt_mdn)) + 
  geom_histogram(bins = 40, color = "black", fill = "cadetblue4") +
  labs(title = "Distribution of Median Student Debt at R1s", 
       subtitle = "", 
       x = "Median Debt (in American dollars)", 
       y = "Frequency") 
debt_hist1
```

```{r}
summary(finR2$grad_debt_mdn)
debt_hist1 <- ggplot(finR2, aes(x = grad_debt_mdn)) + 
  geom_histogram(bins = 40, color = "black", fill = "cadetblue3") +
  labs(title = "Distribution of Median Student Debt at R2s", 
       subtitle = "", 
       x = "Median Debt (in American dollars)", 
       y = "Frequency") 
debt_hist1
```

**Unemployment**

```{r}

print('Unemployment Rate')
summary(finR1$unemp_rate)
unemp_hist1 <- ggplot(finR1, aes(x = unemp_rate)) + 
  geom_histogram(bins = 40, color = "black", fill = "darkseagreen4") +
  labs(title = "Distribution of Post-Graduation Unemployment rates at R1s", 
       subtitle = "", 
       x = "Unemployment Rate", 
       y = "Frequency") 
unemp_hist1
```

```{r}

summary(finR2$unemp_rate)
unemp_hist2 <- ggplot(finR2, aes(x = unemp_rate)) + 
  geom_histogram(bins = 40, color = "black", fill = "darkseagreen3") +
  labs(title = "Distribution of Post-Graduation Unemployment rates at R2s", 
       subtitle = "", 
       x = "Unemployment Rate", 
       y = "Frequency") 
unemp_hist2

```

### R1s vs. R2s

Preliminary normality analysis of the median graduate debt distribution revealed levels of skewness that forbade the use of parametric comparison tests. A Wilcoxon signed-rank test was conducted in order to investigate the differences in median graduate debt between R1 and R2 institutions. Results indicate that median graduate debt significantly differed between R1 and R2 institutions (W = 982, p < 0.01). While a difference in student debt was expected due to the aforementioned price of R1 institution tuition, descriptive statistics revealed graduates of R1 institutions graduated with less average debt (M = 19,652, SD = 3,750) than students from R2 institutions (M = 22,719, SD = 3,785).  

```{r}

wilcox.test(formula=finR$grad_debt_mdn~finR$r_status)

```

In order to further explore student outcomes, the investigators turned their attention to the graduate unemployment rate. The distribution for unemployment exhibited skewness and kurtosis values within acceptable bounds so a Welch’s two-sample t-test was conducted in order to find whether there were significant differences in unemployment rates between R1 and R2 graduates. The test revealed significant differences in rates of unemployment between R1 and R2 graduates t(118.36) = -2.6055, p = 0.01, 95%C.I. [-0.44, -0.06]. On average, the unemployment rate for R1 graduates (M = 3.293, SD = 0.51) was lower than that observed for R2 graduates (M = 3.545, SD = 0.83) 

```{r}

t.test(finR$unemp_rate~finR$r_status)

```

### Regression model  

The racial diversity proportion of an institution was examined as a potential moderator of the relationship between institution type (R1 vs. R2) and unemployment rate. In the first step of the regression analysis institution type was entered into the model and accounted for  5% of the variance in unemployment rate, which was significant, R2= 0.52, F(1,122) = 6.82, p = 0.01.  In the second step of the regression analysis the interaction effect between institution type and racial diversity proportion were entered into the model and explained an additional 4% of the variance in unemployment rate, which was significant, R2= 0.04, F(1,121) = 8.57, p < 0.05. Racial diversity proportion was entered into the model last and accounted for 24% of the variance in unemployment rate, which was significant, R2= 0.24, F(1,122) = 38.8 , p < 0.01. Thus, racial diversity proportion was a significant moderator of the relationship between institution type and unemployment rate.  

**Model step 1**
: TEXT

```{r}

predictor <- finR$r_status
outcome <- finR$unemp_rate

simplelm <- lm(formula = outcome ~ predictor)

```

**Model step 2**
: TEXT

```{r}

predictor <- demR$white
outcome <- finR$unemp_rate

moderatorlm <- lm(formula = outcome ~ predictor)

```

**Model step 3**
: TEXT 

```{r}

predictor <- as.integer(finR$r_status)
moderator <- demR$white
interaction <- predictor * moderator
outcome <- finR$unemp_rate

interactionlm <- lm(formula = outcome ~ interaction)

```

**Complete model**
: TEXT

```{r}

predictor <- as.integer(finR$r_status)
moderator <- demR$white
interaction <- predictor * moderator
outcome <- finR$unemp_rate

wlm <- lm(formula = outcome ~ predictor + moderator + interaction)

```
