# Analysis {#analysis}

## Import Packages
```{r, message=FALSE}
library(moments)
library(dplyr)
library(tidyr)
library(stats)
library(sjstats)
library(ggplot2)
```

```{r}
fin <- read.csv('Proj1Data/cleanFinancials1519.csv')
dem <- read.csv('Proj1Data/cleanDemographics1519.csv')
fin$r_status <- factor(fin$r_status, levels = c("R1", "R2"), labels = c(1,2))
dem$r_status <- factor(dem$r_status, levels = c("R1", "R2"), labels = c(1,2))
```

## Transform Datasets
```{r}
finR <- fin[fin$year == "latest",]
finR1 <- fin[fin$r_status == 1 & fin$year == "latest",]
finR2 <- fin[fin$r_status == 2 & fin$year == "latest",]
demR <- dem[dem$year == "latest",]
demR1 <- dem[dem$r_status == 1 & dem$year == "latest",]
demR2 <- dem[dem$r_status == 2 & dem$year == "latest",]
finR <- subset(finR, university %in% demR$university)
```

## Descriptives

```{r}
# Demographics
summary(demR1$white)
skewness(demR1$white)
kurtosis(demR1$white)
```

```{r}
summary(demR2$white)
skewness(demR2$white)
kurtosis(demR2$white)
```

```{r}
# Cost
print('Tuition cost')
summary(finR1$costt4_a)
cost_hist1 <- ggplot(finR1, aes(x =costt4_a)) + 
  geom_histogram(bins = 40, color = "black", fill = "red") +
  labs(title = "Distribution of Tuition Costs at R1s", 
       subtitle = "", 
       x = "Cost (in American dollars)", 
       y = "Frequency") 
cost_hist1
```

```{r}
summary(finR2$costt4_a)
cost_hist2 <- ggplot(finR2, aes(x =costt4_a)) + 
  geom_histogram(bins = 40, color = "black", fill = "pink") +
  labs(title = "Distribution of Tuition Costs at R2s", 
       subtitle = "", 
       x = "Cost (in American dollars)", 
       y = "Frequency") 
cost_hist2
```

```{r}
# Students debt
print('Student debt')
summary(finR1$grad_debt_mdn)
debt_hist1 <- ggplot(finR1, aes(x = grad_debt_mdn)) + 
  geom_histogram(bins = 40, color = "black", fill = "cadetblue4") +
  labs(title = "Distribution of Median Student Debt at R1s", 
       subtitle = "", 
       x = "Median Debt (in American dollars)", 
       y = "Frequency") 
debt_hist1
```

```{r}
summary(finR2$grad_debt_mdn)
debt_hist1 <- ggplot(finR2, aes(x = grad_debt_mdn)) + 
  geom_histogram(bins = 40, color = "black", fill = "cadetblue3") +
  labs(title = "Distribution of Median Student Debt at R2s", 
       subtitle = "", 
       x = "Median Debt (in American dollars)", 
       y = "Frequency") 
debt_hist1
```

```{r}
# Unemployment
print('Unemployment Rate')
summary(finR1$unemp_rate)
unemp_hist1 <- ggplot(finR1, aes(x = unemp_rate)) + 
  geom_histogram(bins = 40, color = "black", fill = "darkseagreen4") +
  labs(title = "Distribution of Post-Graduation Unemployment rates at R1s", 
       subtitle = "", 
       x = "Unemployment Rate", 
       y = "Frequency") 
unemp_hist1
```

```{r}
summary(finR2$unemp_rate)
unemp_hist2 <- ggplot(finR2, aes(x = unemp_rate)) + 
  geom_histogram(bins = 40, color = "black", fill = "darkseagreen3") +
  labs(title = "Distribution of Post-Graduation Unemployment rates at R2s", 
       subtitle = "", 
       x = "Unemployment Rate", 
       y = "Frequency") 
unemp_hist2
```

## Wilcox signed-rank test  
Preliminary normality analysis of the median graduate debt distribution revealed levels of skewness that forbade the use of parametric comparison tests. A Wilcoxon signed-rank test was conducted in order to investigate the differences in median graduate debt between R1 and R2 institutions. Results indicate that median graduate debt significantly differed between R1 and R2 institutions (W = 982, p < 0.01). While a difference in student debt was expected due to the aforementioned price of R1 institution tuition, descriptive statistics revealed graduates of R1 institutions graduated with less average debt (M = 19,652, SD = 3,750) than students from R2 institutions (M = 22,719, SD = 3,785).  
```{r}
wilcox.test(formula=finR$grad_debt_mdn~finR$r_status)
```

## Welch's two-sample t-test  
In order to further explore student outcomes, the investigators turned their attention to the graduate unemployment rate. The distribution for unemployment exhibited skewness and kurtosis values within acceptable bounds so a Welch’s two-sample t-test was conducted in order to find whether there were significant differences in unemployment rates between R1 and R2 graduates. The test revealed significant differences in rates of unemployment between R1 and R2 graduates t(118.36) = -2.6055, p = 0.01, 95%C.I. [-0.44, -0.06]. On average, the unemployment rate for R1 graduates (M = 3.293, SD = 0.51) was lower than that observed for R2 graduates (M = 3.545, SD = 0.83)  
```{r}
t.test(finR$unemp_rate~finR$r_status)
```

## Regression model  

The racial diversity proportion of an institution was examined as a potential moderator of the relationship between institution type (R1 vs. R2) and unemployment rate. In the first step of the regression analysis institution type was entered into the model and accounted for  5% of the variance in unemployment rate, which was significant, R2= 0.52, F(1,122) = 6.82, p = 0.01.  In the second step of the regression analysis the interaction effect between institution type and racial diversity proportion were entered into the model and explained an additional 4% of the variance in unemployment rate, which was significant, R2= 0.04, F(1,121) = 8.57, p < 0.05. Racial diversity proportion was entered into the model last and accounted for 24% of the variance in unemployment rate, which was significant, R2= 0.24, F(1,122) = 38.8 , p < 0.01. Thus, racial diversity proportion was a significant moderator of the relationship between institution type and unemployment rate.  

## Model step 1  
```{r}
predictor <- finR$r_status
outcome <- finR$unemp_rate

simplelm <- lm(formula = outcome ~ predictor)
```

## Model step 2  
```{r}
predictor <- demR$white
outcome <- finR$unemp_rate

moderatorlm <- lm(formula = outcome ~ predictor)
```

## Model step 3  
```{r}
predictor <- as.integer(finR$r_status)
moderator <- demR$white
interaction <- predictor * moderator
outcome <- finR$unemp_rate

interactionlm <- lm(formula = outcome ~ interaction)
```

## Complete model
```{r}
predictor <- as.integer(finR$r_status)
moderator <- demR$white
interaction <- predictor * moderator
outcome <- finR$unemp_rate

wlm <- lm(formula = outcome ~ predictor + moderator + interaction)
```
